---
title: "Assessing Home Price Determinants: A Regression Analysis of Ames Housing Data"
output:
  html_document:
    df_print: paged
  pdf_document: default
author: 
-   Ryan Croce
-   Keely Kinnane
-   Sai Teja Yapuram
-   Kavindu Wellalage
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Narrative

There are many factors ranging from location, material quality, and property size, among others, that may influence the sales price of a home. We aim to better understand how all these factors drive market prices in order to determine which features are relevant, given their relationship to sales price, as well as their collective impact on the accuracy of our predictions. The Ames dataset provides a comprehensive selection of these property features, which we can use to explore these questions. Our goal is to construct a model to analyze the selected parameters and assess their impact on price and to make predictions for different types of homes. We hypothesize that the variables `LotArea`, `Neighborhood`, `OverallQual`, and `OverallCond` are the most relevant predictors for the response variable `SalePrice`. We anticipate that a model comprised solely of these variables will lead to the smallest Root Mean Square Error (RMSE), thus indicating the most accurate predictions of home sale prices. Our results can be used to assist home sellers, policymakers, and real estate professionals in better identifying the most important factors influencing home prices.

# Dataset Description

The Ames Housing Data set curated by Dean De Cock from Truman State University. The original data set describes the sale of individual residential property in Ames, Iowa from 2006 to 2010. The original data set contains 2930 observations and a large number of explanatory
variables (23 nominal, 23 ordinal, 14 discrete, and 20 continuous) involved in assessing home values. The version used for this project is a subset of the original data set that contains 766 homes and the 28 variables. The subset is selected by following the steps given below.

-   Removed large houses with an above ground living area of more than 1500 square feet.

-   Filtered the data set to include houses that had a 'Normal' sales condition.

-   Included the top five neighborhoods with the most houses for sale.

-   Removed rows with `NA` values.

-   Renamed a few variables to make their meaning clearer.

# Data Dictionary

The full data set contains 766 homes and the following 28 variables:

-   `LotArea`: Lot size in square feet. A numeric variable.
-   `Neighborhood`: Physical location within Ames city limits. A categorical variable with the following levels:
    -   "NAmes" : North Ames
    -   "Sawyer" : Sawyer
    -   "OldTown" : Old Town
    -   "Edwards" : Edwards
    -   "CollgCr" : College Creek
-   `OverallQual`: Overall quality of the house's material and finish. The scale ranges from 1 (Very Poor) to 9 (Very Excellent). A numeric variable.
-   `OverallCond`: Overall condition rating. The scale ranges from 1 (Very Poor) to 9 (Very Excellent). A numeric variable.
-   `YearBuilt`: Original construction date. A numeric variable.
-   `YearRemodAdd`: Remodel date. A numeric variable.
-   `BsmtFinSf1`: Type 1 finished square feet. A numeric variable.
-   `BsmtFinSf2`: Type 2 finished square feet. A numeric variable.
-   `TotalBsmtSf`: Total square feet of basement area. A numeric variable.
-   `FirstFlrSf`: First floor square feet. A numeric variable.
-   `SecondFlrSf`: Second floor square feet. A numeric variable.
-   `GrLivArea`: Above grade (ground) living area square feet. A numeric variable.
-   `BsmtFullBath`: Number of full bathrooms in the basement. A numeric variable.
-   `BsmtHalfBath`: Number of half baths in the basement. A numeric variable.
-   `FullBath`: Number of full bathrooms above ground. A numeric variable.
-   `HalfBath`: Number of half baths above ground. A numeric variable.
-   `BedroomAbvGr`: Number of Bedrooms above ground. A numeric variable.
-   `KitchenAbvGr`: Number of Kitchens above ground. A numeric variable.
-   `TotRmsAbvGrd`: Total rooms above ground (does not include bathrooms). A numeric variable.
-   `Fireplaces`: Number of fireplaces. A numeric variable. A numeric variable.
-   `GarageCars`: Size of garage in car capacity. A numeric variable.
-   `GarageArea`: Size of garage in square feet. A numeric variable.
-   `WoodDeckSf`: Wood deck area in square feet. A numeric variable.
-   `OpenPorchSf`: Open porch area in square feet. A numeric variable.
-   `EnclosedPorch`: Enclosed porch area in square feet. A numeric variable.
-   `ThreeSsnProch`: Three season porch area in square feet. A numeric variable.
-   `ScreenPorch`: Screen porch area in square feet. A numeric variable.
-   `SalePrice`: The property's sale price in dollars. A numeric response variable.

# Evidence of Data

```{r, echo=FALSE}
data = read.csv('ames_housing.csv')
head(data, nrow=10)
```
