---
title: "Sale Price Prediction of Ames Houses"
output:
  pdf_document: default
  html_document:
    df_print: paged
author: 
-   Ryan
-   Keely Kinnane
-   Sai Teja Yapuram
-   Kavindu Wellalage
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Dictionary

The full data set contains 766 homes and the following 28 variables:

-   `LotArea`: Lot size in square feet. A numeric variable.
-   `Neighborhood`: Physical location within Ames city limits. A categorical variable with the following levels:
    -   "NAmes" : North Ames
    -   "Sawyer" : Sawyer
    -   "OldTown" : Old Town
    -   "Edwards" : Edwards
    -   "CollgCr" : College Creek
-   `OverallQual`: Overall quality of the house's material and finish. The scale ranges from 1 (Very Poor) to 9 (Very Excellent). A numeric variable.
-   `OverallCond`: Overall condition rating. The scale ranges from 1 (Very Poor) to 9 (Very Excellent). A numeric variable.
-   `YearBuilt`: Original construction date. A numeric variable.
-   `YearRemodAdd`: Remodel date. A numeric variable.
-   `BsmtFinSf1`: Type 1 finished square feet. A numeric variable.
-   `BsmtFinSf2`: Type 2 finished square feet. A numeric variable.
-   `TotalBsmtSf`: Total square feet of basement area. A numeric variable.
-   `FirstFlrSf`: First floor square feet. A numeric variable.
-   `SecondFlrSf`: Second floor square feet. A numeric variable.
-   `GrLivArea`: Above grade (ground) living area square feet. A numeric variable.
-   `BsmtFullBath`: Number of full bathrooms in the basement. A numeric variable.
-   `BsmtHalfBath`: Number of half baths in the basement. A numeric variable.
-   `FullBath`: Number of full bathrooms above ground. A numeric variable.
-   `HalfBath`: Number of half baths above ground. A numeric variable.
-   `BedroomAbvGr`: Number of Bedrooms above ground. A numeric variable.
-   `KitchenAbvGr`: Number of Kitchens above ground. A numeric variable.
-   `TotRmsAbvGrd`: Total rooms above ground (does not include bathrooms). A numeric variable.
-   `Fireplaces`: Number of fireplaces. A numeric variable. A numeric variable.
-   `GarageCars`: Size of garage in car capacity. A numeric variable.
-   `GarageArea`: Size of garage in square feet. A numeric variable.
-   `WoodDeckSf`: Wood deck area in square feet. A numeric variable.
-   `OpenPorchSf`: Open porch area in square feet. A numeric variable.
-   `EnclosedPorch`: Enclosed porch area in square feet. A numeric variable.
-   `ThreeSsnProch`: Three season porch area in square feet. A numeric variable.
-   `ScreenPorch`: Screen porch area in square feet. A numeric variable.
-   `SalePrice`: The property's sale price in dollars. A numeric response variable.

# Evidence of Data

```{r}
data = read.csv('ames_housing.csv')
head(data, nrow=10)
```
